Concurrent Programming in Harmony

### 2021-12-24 book/paper.tex
Pass over Chapter 9 finished

Method <{incrementer}> takes a parameter called <{self}>.
It increments <{count}> and sets <{done[self]}> to <{True}>.
It then waits until the other thread is done.
(<{await c}> is shorthand for <{while not c: pass}>.)
After that, method <{incrementer}>
verifies that the value of <{count}> equals~2.
% Like Python, Harmony supports \textbf{assert} statements with two arguments:
% the first is a Boolean condition, and the second is a value that is
% reported if the condition fails, in this case
% the value of <{count}>.

@@

To support implementing locks,
many CPUs have an atomic ``test-and-set'' (TAS)
\index{test-and-set}%
\index{TAS}%
operation.
Method \texttt{test\_and\_set} in \autoref{fig:tas} shows its specification.
Here <{s}> points to a shared Boolean variable and <{p}>
to a private Boolean variable, belonging to some thread.
The operation copies the value of the shared variable to the
private variable (the ``test'')
and then sets the shared variable to <{True}> (``set'').

\autoref{fig:tas} goes on to implement mutual exclusion for
a set of \texttt{N} threads.
The approach is called \emph{spinlock},
\index{spinlock}%
because each thread is ``spinning'' (executing a tight loop) until
it can acquire the lock.
The program uses $\mathtt{N}+1$ variables.
Variable <{shared}> is initialized to <{False}> while
\textit{private}[<{i}>] for each thread <{i}> is initialized to <{True}>.

An important invariant, $\mathcal{I}_1$, of the program is that at any time at most
one of these variables is <{False}>.
Another invariant, $\mathcal{I}_2(i)$, is that if thread <{i}> is executing between
lines $18 \cdots 22$ (which includes the critical section),
then \textit{private}[<{i}>] = <{False}>.
Between the two (i.e., $\mathcal{I}_1 \land \forall i: \mathcal{I}_2(i)$),
it is clear that only one thread can be in the
critical section at the same time.
For those who read through \autoref{ch:peterson}
$\mathcal{I}_1 \land \forall i: \mathcal{I}_2(i)$ is an inductive invariant.

@@

The Boolean type has only two possible values: <{False}> and
<{True}>.  Unlike in Python, in Harmony booleans are distinct
from integers, and in particular $\mathbf{False} < 0$.  In statements
and expressions where booleans are expected, it is not possible to
substibute values of other types.

>>>>

Method <{incrementer}> takes a parameter called <{self}>.
It increments <{count}> and sets <{done[self]}> to <{True}>.
It then waits until the other thread is done.
(<{await c}> is shorthand for <{while not c: pass}>.)
After that, method <{incrementer}>
verifies that the value of <{count}> equals~2.

@@

To support implementing locks,
many CPUs have an atomic ``test-and-set'' (TAS)
\index{test-and-set}%
\index{TAS}%
operation.
Method \texttt{test\_and\_set} in \autoref{fig:tas} shows its specification.
Here <{s}> points to a shared boolean variable and <{p}>
to a private boolean variable, belonging to some thread.
The operation copies the value of the shared variable to the
private variable (the ``test'')
and then sets the shared variable to <{True}> (``set'').

\autoref{fig:tas} goes on to implement mutual exclusion for
a set of \texttt{N} threads.
The approach is called \emph{spinlock},
\index{spinlock}%
because each thread is ``spinning'' (executing a tight loop) until
it can acquire the lock.
The program uses $\mathtt{N}+1$ boolean variables.
Variable <{shared}> is initialized to <{False}> while
\textit{private}[<{i}>] for each thread <{i}> is initialized to <{True}>.

An important invariant, $\mathcal{I}_1$, of the program is that at any time at most
one of these variables is <{False}>.
Another invariant, $\mathcal{I}_2(i)$, is that if thread <{i}> is
in the critical section,
then \textit{private}[<{i}>] = <{False}>.
Between the two (i.e., $\mathcal{I}_1 \land \forall i: \mathcal{I}_2(i)$),
it is clear that only one thread can be in the
critical section at the same time.
% For those who read through \autoref{ch:peterson}
% $\mathcal{I}_1 \land \forall i: \mathcal{I}_2(i)$ is an inductive invariant.

@@

The boolean type has only two possible values: <{False}> and
<{True}>.  Unlike in Python, in Harmony booleans are distinct
from integers, and in particular $\mathbf{False} < 0$.  In statements
and expressions where booleans are expected, it is not possible to
substibute values of other types.
